That’s a Big-Ass Chicken: A Mixed Methods Examination of Urban Backyard Chicken Husbandry

### 2025-03-07 BAC_Manuscript_Final.qmd
desperately trying to push through a corrected Qmd that will now render to pdf

---
title: "That’s a Big-Ass Chicken: A Mixed Methods Examination of Urban Backyard Chicken Husbandry"
shorttitle: "That's a Big-Ass Chicken"
author:
  - name: Karen Veronica Becerra
    corresponding: true
    orcid: 0009-0006-4967-0955
    email: kvbecerra@Uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
        id: uchicago
    role:
      - conceptualization
      - formal anlaysis
      - methodology
      - project administration
      - software
      - visualization
      - editing 
  - name: Kristen M. Johnson
    orcid: 0009-0008-3757-5795
    email: kmjohnson25@uchicago.edu 
    affiliations:
      - ref: uchicago
    role:
      - conceptualization
      - data curation
      - methodology
      - project administration
      - software
      - writing
      - editing
  - name: Michael Moroz
    orcid: 0009-0000-9772-0001
    email: mrmoroz@uchicago.edu
    affiliations:
      - ref: uchicago
    role:
      - formal anlaysis
      - software
      - visualization
      - editing
abstract: "This study investigates the growing phenomenon of backyard chicken keeping in urban and suburban environments, a practice that has gained significant momentum amid rising commercial egg prices. Employing a mixed methods approach, we analyzed quantitative data from ______. Concurrently, we conducted thematic analysis of in-depth interviews with backyard chicken owners to explore perceived benefits and challenges. Our findings indicate that while initial setup costs average $487, households recover this investment within 14 months through egg production, with additional benefits including reduced food waste, improved garden soil quality, and reported increases in household well-being. Challenges identified include time commitment (averaging 28 minutes daily), navigating municipal regulations, and managing seasonal variations in productivity. Factor analysis revealed that success factors cluster around three dimensions: appropriate infrastructure, consistent maintenance routines, and community support networks. Ultimately, our results support the hypothesis that the benefits of backyard chicken keeping outweigh the challenges for most practitioners, though successful implementation depends on household-specific variables and community context. This research contributes to growing literature on urban agriculture while providing practical insights for potential chicken keepers, municipal policymakers, and sustainability advocates."
keywords: [chickens, backyard, eggs, pros and cons, bigger is better]
author-note:
  disclosures:
    conflict of interest: The authors have no conflicts of interest to declare.
  financial-support: "This study was supported by a grant from The International Society for Fowl Language."
  gratitude: "The authors are grateful to Eddy the Chicken for her endless inspiration and reluctant support."
impact-statement: "This study illustrates not only how, but also why one should raise big-ass backyard chickens. The results may be used by numerous individuals to convince their domestic partners (such as the second author's husband) to build a backyard coop and give in to the siren squawk of that sweet sweet chicken-raisin' lifestyle."
bibliography: Chickens.bib
format: 
  #apaquarto-docx: default 
  #apaquarto-html: default    # remember to reverse commenting out other formats when turning project in
  apaquarto-pdf: default
floatsintext: true
documentmode: man
donotrepeattitle: true
mainfont: "Times New Roman"
fontsize: 12pt

@@

```{r, echo=FALSE, include=FALSE}

# Load necessary libraries
library(readr)
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(flextable)

@@

[paragraph about why bigger is better, with regards to egg production and everything else]

Previous studies have largely focused either on large-scale commercial poultry operations or on historical agricultural practices [@bistSustainablePoultryFarming2024; @grzinicIntensivePoultryFarming2023; @jeniOverviewHealthChallenges2021].  While proponents extol the virtues of fresh eggs and animated garden assistants, detractors cite concerns ranging from zoonotic disease transmission to neighborhood noise pollution---a concern that anyone who has experienced a rooster's enthusiasm for announcing dawn can readily appreciate. Surprisingly, of those studies that do consider the costs and benefits of backyard chicken keeping, none that to the authors knowledge look into the phenomenon of the Big-Ass Chicken (BAC)[@ayalaReviewPathogenTransmission2020; @fischerDefenceBackyardChickens2019; @schindlerBackyardChickensFrontYard; @singhBackyardPoultryFarming2022]. This study aims to fill that obvious (and sizeable) gap.

@@

# Method
## Participants
### Chickens
We looked at 50 Gallus gallus domesticus chicks, starting from hatching...

@@

## Procedure
### Quantitative
We measured the weight of chickens ....

@@

# Results
## Chickens  
```{r, echo=FALSE, include=FALSE}

#| label: data-analysis

@@

```{r, echo=FALSE}
## Making the first plot (weight on the y-axis, time on the x-axis, line plots grouped by chicken ID (meaning each chicken will have its own colour on the line plot))

@@

```{r, echo=FALSE, include=FALSE}
## Ok now we have visualized the different growth trends of the various chickens. Now we probably want to look at how diet might factor into this trend. First let's calculate the mean growth rates of all the various chickens according to diet (rated 1 to 4)

@@

```{r, echo=FALSE}
## Ok so now we see that diet 3 has chickens with the highest mean growth rates (per day). Diet 4 comes in second, then diet 2, then diet 1. Let's visualize everything in a ggplot.

@@

```{r, echo=FALSE, include=FALSE}
## So if we are to infer that diet is a good predictor of growth rates (obviously in addition to time), we should see diet as a significant predictor of weight in a linear regression. Let's try that now.

@@

```{r}
#| echo: FALSE
#| label: regression-plot
#| fig-cap: Regression of Weight on Time by Diet

@@

## Eggs 
#| echo: FALSE
#| label: regression-plot
#| fig-cap: Regression of Weight on Time by Diet

## Eggs 
```{r, echo=FALSE, include=FALSE}
# Load necessary library
set.seed(123)  # Created fake data because we want Kristen to get chickens

# Load necessary library
set.seed(123)

@@

```{r, echo=FALSE, include=FALSE}
# Load necessary libraries
library(knitr)
library(kableExtra)

# Format the table with kable
chicken_data %>%
  kable("html", caption = "Chicken Data: Egg Size and Weight by Breed") %>%
  kable_styling(
    font_size = 12,  # Change font size to match APA formatting
    full_width = FALSE
  ) %>%
  column_spec(1, bold = TRUE)  # Bold the first column (Chicken_ID) for emphasis
```

```{r, echo=FALSE, include=FALSE}

@@

```{r, echo=FALSE}
# Run the correlation between Weight and Eggs_Laid (Pearson correlation)
correlation_result <- cor.test(chicken_data$Weight, chicken_data$Eggs_Laid)

@@

```{r, echo=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)

@@

###Practical Challenges and Considerations

>>>>

---
title: "That’s a Big-Ass Chicken: A Mixed Methods Examination of Urban Backyard Chicken Husbandry"
shorttitle: "That's a Big-Ass Chicken"
author:
  - name: Karen Veronica Becerra
    corresponding: true
    orcid: 0009-0006-4967-0955
    email: kvbecerra@Uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
        id: uchicago
    role:
      - conceptualization
      - formal anlaysis
      - methodology
      - project administration
      - software
      - visualization
      - editing 
  - name: Kristen M. Johnson
    orcid: 0009-0008-3757-5795
    email: kmjohnson25@uchicago.edu 
    affiliations:
      - ref: uchicago
    role:
      - conceptualization
      - data curation
      - formal analysis
      - methodology
      - project administration
      - software
      - writing
      - editing
  - name: Michael Moroz
    orcid: 0009-0000-9772-0001
    email: mrmoroz@uchicago.edu
    affiliations:
      - ref: uchicago
    role:
      - formal anlaysis
      - software
      - visualization
      - editing
abstract: "This study investigates the growing phenomenon of backyard chicken keeping in urban and suburban environments, a practice that has gained significant momentum amid rising commercial egg prices. Employing a mixed methods approach, we analyzed quantitative data from ______. Concurrently, we conducted thematic analysis of in-depth interviews with backyard chicken owners to explore perceived benefits and challenges. Our findings indicate that while initial setup costs average $487, households recover this investment within 14 months through egg production, with additional benefits including reduced food waste, improved garden soil quality, and reported increases in household well-being. Challenges identified include time commitment (averaging 28 minutes daily), navigating municipal regulations, and managing seasonal variations in productivity. Factor analysis revealed that success factors cluster around three dimensions: appropriate infrastructure, consistent maintenance routines, and community support networks. Ultimately, our results support the hypothesis that the benefits of backyard chicken keeping outweigh the challenges for most practitioners, though successful implementation depends on household-specific variables and community context. This research contributes to growing literature on urban agriculture while providing practical insights for potential chicken keepers, municipal policymakers, and sustainability advocates."
keywords: [chickens, backyard, eggs, pros and cons, bigger is better]
author-note:
  disclosures:
    conflict of interest: The authors have no conflicts of interest to declare.
  financial-support: "This study was supported by a grant from The International Society for Fowl Language."
  gratitude: "The authors are grateful to Eddy the Chicken for her endless inspiration and reluctant support."
impact-statement: "This study illustrates not only how, but also why one should raise big-ass backyard chickens. The results may be used by numerous individuals to convince their domestic partners (such as the second author's husband) to build a backyard coop and give in to the siren squawk of that sweet sweet chicken-raisin' lifestyle."
bibliography: Chickens.bib
format: 
  #apaquarto-docx: default 
  apaquarto-html: default    # remember to reverse commenting out other formats when turning project in
  apaquarto-pdf: default
floatsintext: true
documentmode: man
donotrepeattitle: true
mainfont: "Times New Roman"
fontsize: 12pt

@@

```{r}
#| label: Setup
#| echo: FALSE 
#| include: FALSE

# Load necessary libraries
library(readr)
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(flextable)
library(knitr)
library(kableExtra)
library(dplyr)
```

@@

[paragraph about why bigger is better, with regards to egg production and everything else^[This paragraph did not end up getting written because really what more do we need to say.]]

Previous studies have largely focused either on large-scale commercial poultry operations or on historical agricultural practices [@bistSustainablePoultryFarming2024; @grzinicIntensivePoultryFarming2023; @jeniOverviewHealthChallenges2021].  While proponents extol the virtues of fresh eggs and animated garden assistants, detractors cite concerns ranging from zoonotic disease transmission to neighborhood noise pollution---a concern that anyone who has experienced a rooster's enthusiasm for announcing dawn can readily appreciate. Surprisingly, of those studies that do consider the costs and benefits of backyard chicken keeping, none that to the authors knowledge look into the phenomenon of the **Big-Ass Chicken** (BAC)[@ayalaReviewPathogenTransmission2020; @fischerDefenceBackyardChickens2019; @schindlerBackyardChickensFrontYard; @singhBackyardPoultryFarming2022]. This study aims to fill that obvious (and sizeable) gap.

@@

# Method
## Participants
### Chickens
We looked at 50 Gallus gallus domesticus chicks, starting from hatching.

@@

## Procedure
### Quantitative
We measured the body weights of the chickensat birth and every second day thereafter until day 20. They were also measured on day 21. There were four groups of chicks on different protein diets to test the effect of protein in 4 types of feed on early development of chicks. The 4 feeds are a control (normal diet) and 3 test feeds (10%, 20%, and 40% protein replacement).

@@

# Results
## Chickens  
```{r}
#| label: data-analysis_setup_chicken-weight
#| echo: FALSE 
#| include: FALSE

@@

```{r}
#| label: chicken_weight_growth_plot
#| echo: FALSE 
#| include: FALSE

## Making the first plot (weight on the y-axis, time on the x-axis, line plots grouped by chicken ID (meaning each chicken will have its own colour on the line plot))

@@

```{r}
#| label: chicken_weight_growth-rate
#| echo: FALSE 
#| include: FALSE

## Ok now we have visualized the different growth trends of the various chickens. Now we probably want to look at how diet might factor into this trend. First let's calculate the mean growth rates of all the various chickens according to diet (rated 1 to 4)

@@

```{r}
#| label: chicken_weight_growth-rate_plot
#| echo: FALSE 
#| include: FALSE

## Ok so now we see that diet 3 has chickens with the highest mean growth rates (per day). Diet 4 comes in second, then diet 2, then diet 1. Let's visualize everything in a ggplot.

@@

```{r}
#| label: chicken_weight_regression
#| echo: FALSE 
#| include: FALSE

## So if we are to infer that diet is a good predictor of growth rates (obviously in addition to time), we should see diet as a significant predictor of weight in a linear regression. Let's try that now.

@@

```{r}
#| label: chicken_weight_regression-plot
#| echo: FALSE
#| fig-cap: Regression of Weight on Time by Diet

@@

## Eggs 
```{r}
#| label: chicken_egg_data
#| echo: FALSE 
#| include: FALSE

# Load necessary library
set.seed(123)  # Created fake data because we want Kristen to get chickens

@@

```{r}
#| label: chicken_egg_size_weight
#| echo: FALSE 
#| include: FALSE

# Format the table with kable
chicken_data %>%
  kable(caption = "Chicken Data: Egg Size and Weight by Breed") %>%
  kable_styling(
    font_size = 12,  # Change font size to match APA formatting
    full_width = FALSE
  ) %>%
  column_spec(1, bold = TRUE)  # Bold the first column (Chicken_ID) for emphasis
```

```{r}
#| label: chicken-egg_summary_stats
#| echo: FALSE
#| include: FALSE

@@

```{r}
#| label: chicken-egg_correlation
#| echo: FALSE
#| include: FALSE

# Run the correlation between Weight and Eggs_Laid (Pearson correlation)
correlation_result <- cor.test(chicken_data$Weight, chicken_data$Eggs_Laid)

@@

```{r}
#| label: chicken-egg_correlation_plot
#| echo: FALSE

@@

### Practical Challenges and Considerations
